---
title: "Data Import and Tidy"
author: "Rose Killian"
date: "3/6/2023"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

##### Load libraries and settings

```{r libraries and settings, warning=FALSE, message=FALSE}

library(tidyverse)
library(viridis)
library(readxl)
library(lubridate)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##### Version notes
Discrepancy with IDs 299 and 300 resolved. Both are aliases for ID 10. These observations can now be assigned to their group and will no longer be dropped. 

I was also given data from REDcap that lists the dates of each patient visit and the week that that visit fell on. 

### Import

#### ASA24 Data
```{r import asa24}
asa24_df = 
  read_excel(
    "data/TOTAL ASA24 DATA - Luis healthy and unhealthy foods.xlsx",
    sheet = "HelmCroh + HelmCr") %>%
  janitor::clean_names() %>% 
  select(
    user_name, recall_no, intake_start_date_time, intake_end_date_time, reporting_date, f_total, v_total, v_legumes, g_whole, pf_poult, pf_seafd_hi, pf_seafd_low, pf_nutsds, pf_legumes, d_yogurt, g_refined, pf_meat, pf_curedmeat, solid_fats, add_sugars, kcal, prot, tfat, carb, sugr, fibe, sfat, chole
  ) %>% 
mutate(
  id = str_replace(user_name, "HelmCroh", ""))

#R gets mad if I try to remove the different username prefixes in the same step. I'm sure there is a more elegant way to do this but this works so for now this is how it is. Code above imports, cleans names, and removes the first prefix used for user_name and makes new variable "id". Code below removes second prefix and converts vector to numeric.

asa24_df = 
  asa24_df %>% 
  mutate(id = str_replace(id, "HelmCr", "")) %>% 
  mutate(
    id = as.numeric(id)) %>% 
  relocate(id) %>% 
  arrange(id, reporting_date)
```

#### Visits data
```{r visit dates}
visits_df = 
  read_excel(
    "data/AHolisticDietInterve_DATA_2023-03-06_1029.xlsx") %>% 
  janitor::clean_names() %>% 
  mutate(
    group = as_factor(registration_arm_1_ramdom),
    group = recode_factor(group, `1` = "Group 1", `2` = "Group 2", `3` = "Group 3P", `4` = "Group 3C")) %>% 
  rename(
    `1a` = visit_1_screening_arm_1_asa24_weekdate,
    `1b` = visit_1_screening_arm_1_dateasa24_weekday2,
    `1c` = visit_1_screening_arm_1_asa24_weekenddate,
    `2a` = week_12_ibdzoom_5_arm_1_asa24_weekdate,
    `2b` = week_12_ibdzoom_5_arm_1_dateasa24_weekday2,
    `2c` = week_12_ibdzoom_5_arm_1_asa24_weekenddate,
    `3a` = week_24_ibd_zoom_7_arm_1_asa24_weekdate,
    `3b` = week_24_ibd_zoom_7_arm_1_dateasa24_weekday2,
    `3c` = week_24_ibd_zoom_7_arm_1_asa24_weekenddate,
    `4a` = visit_3_last_ibds_arm_1_asa24_weekdate,
    `4b` = visit_3_last_ibds_arm_1_dateasa24_weekday2,
    `4c` = visit_3_last_ibds_arm_1_asa24_weekenddate,
    id = registration_arm_1_internal_pt_id
  ) %>% 
  select(
    id, group, `1a`, `1b`, `1c`, `2a`, `2b`, `2c`, `3a`, `3b`, `3c`, `4a`, `4b`, `4c`
  ) %>% 
  pivot_longer(`1a`:`4c`,
               names_to = "visit",
               values_to = "reporting_date",
               names_pattern = "(.*)[a-c]") %>% 
  mutate(
    visit = as_factor(visit),
    visit = recode_factor(visit, `1` = "Week 0", `2` = "Week 12", `3` = "Week 24", `4` = "Week 36")
  ) %>% 
  relocate(id, reporting_date) %>% 
  arrange(id, reporting_date) %>% 
  distinct()
```


#### Completion data

```{r import completion}
completion_df = 
  read_csv(
    "data/AHolisticDietInterve_DATA_2023-03-06_1242.csv") %>% 
      janitor::clean_names() %>% 
  rename(id = registration_arm_1_internal_pt_id,
         completedyesno = registration_arm_1_completedyesno)

completion_df %>% 
  group_by(completedyesno) %>% 
  summarise(n_obs = n()) %>% 
  knitr::kable()
```
Looking at the variable for study completion, we have 93 patients that have completed the study, 18 that have been marked as dropped out/incomplete, and 4 who are missing any designation (possibly still in process of completing the study). Patients marked as dropped out will be dropped from the data.

##### Table of Dropouts

```{r dropout table}
completion_df %>% 
  filter(completedyesno == 0) %>% 
  select(id, completedyesno) %>% 
  knitr::kable()
```


```{r test of drop dropouts}
#start_completion =
#  start_completion %>% 
#  filter(is.na(completedyesno) | !completedyesno == 0)

#This code was used to make sure that only the participants marked as having dropped out were dropped from the dataset.
```


### Join Datasets

Joining the data as well as dropping the dropouts noted earlier

```{r merge}
eda_df = 
  left_join(asa24_df, visits_df, by = c("id", "reporting_date"))

eda_df = 
  left_join(eda_df, completion_df, by = "id") %>% 
  filter(is.na(completedyesno) | !completedyesno == 0) %>% 
  relocate(id, reporting_date, group, visit)

```

### Visualization

```{r merged data overview}
skimr::skim(eda_df)
```

148 entries missing their visit date (and group). Need to try matching visit date with intake_start_date-time? 

* Absolutely not. Missing increased to 830!